---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

with SQL, you can filter columns with SELECT and rows with WHERE.  Let's look at an example using the Lahman baseball database.  First, we need to load some packages.

```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```
#--------------------
Suppose we want to see the homerun totals for the 1927 Yankees.  We could write the following:

```{r}
query<-"SELECT playerID,yearID,teamID,HR from Batting
WHERE teamID='NYA'and yearID=1927"
sqldf(query)
```
#---------------------
Finding players with fewer than 60 strikeout but more than 40 homeruns
```{r}
query<-"SELECT playerID, teamID, yearID, HR, SO FROM Batting WHERE SO<60 and HR>40"

sqldf(query)
```
#--------------------
All instances where PHI hit more than 30 HR in the 1970s
```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting WHERE teamID='PHI' and yearID>=1970 and yearID<1980 and HR>30"

sqldf(query)
```

#------------------
##Order By

Find every instance of a player hitting more than 50 HR's, but let's have the players with the most HR at the top of the list.

```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

#-----------------
Find all instance of a player who batted at least 400 times, but struck out less then 10 times with the least amount of strikeouts at the top.
```{r}
query<-"SELECT playerID, teamID, yearID, SO, AB FROM Batting WHERE AB>=400 and SO<10
ORDER BY SO"

sqldf(query)
```

